---
title: "Migration as the Demographic Engine of Industrial Policy: Evidence from Argentina"
subtitle: "Gravity Models and Mediation Analysis"
author:
  - name: "Author Name"
    affiliation: "Institution"
    corresponding: true
date: last-modified
abstract: |
  We examine migration as the primary mechanism through which industrial promotion
  zones reshape regional demographic structures in Argentina (1991–2022). Using
  a PPML gravity framework and origin-destination flow data from five census rounds,
  we decompose interprovincial migration flows into policy-related and secular
  components. Formal mediation analysis reveals that migration accounts for
  approximately [X]% of the total industrial policy effect on local population growth
  identified in Paper 1. Migration is positively skill-selective into treated areas
  and negatively selective in sending regions, contributing to regional human capital
  polarization. These findings highlight migration as both a vehicle and a consequence
  of place-based industrialization.
keywords:
  - migration
  - gravity model
  - mediation analysis
  - industrial policy
  - Argentina
  - origin-destination flows
format:
  pdf:
    documentclass: article
    classoption: [12pt, a4paper]
    number-sections: true
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{amsmath}
        \usepackage{float}
        \usepackage{setspace}
        \doublespacing
  html:
    toc: true
    number-sections: true
    theme: cosmo
bibliography: references.bib
csl: apa.csl
execute:
  echo: false
  warning: false
  message: false
  cache: true
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(targets)
library(here)
library(gt)

od_period     <- targets::tar_read(od_period)
skill_decomp  <- targets::tar_read(skill_decomp)
did_effects   <- targets::tar_read(did_effects)
```

# Introduction {#sec-intro}

Regional population change is proximately the sum of natural increase and net migration.
In the context of working-age policies like industrial promotion — which primarily affect
labor market conditions — migration is theoretically the dominant proximate mechanism
linking economic incentives to population dynamics [@greenwood_1975; @sahota_1968].
Despite this, migration as a **mediator** of place-based policy effects has received
surprisingly limited attention in the development economics literature.

This paper fills that gap by analyzing **whether, how much, and for whom** migration
mediates the population growth effects of Argentina's industrial promotion zones.
We build directly on the causal effect estimates established in Paper 1 of this series.

## Contribution

1. First gravity-model analysis of Argentine interprovincial migration using harmonized
   census microdata spanning five decades.
2. Formal mediation analysis decomposing total industrial policy → population effects
   into migration-mediated and direct components.
3. Evidence on skill-selective sorting induced by place-based policy.

# Migration Theory and Industrial Policy {#sec-theory}

## The Migration Mediation Hypothesis

The Todaro-Harris model predicts migration from lower-wage rural regions to higher-wage
urban-industrial regions when expected wage differentials exceed migration costs
[@harris_todaro_1970]. Industrial promotion, by subsidizing labor demand in
otherwise unattractive regions, can invert this logic — generating in-migration to
regions whose natural comparative advantage would not otherwise attract workers.

We formalize this as a **mediation model**:

$$
\text{Population}_{it} = \alpha + \delta \cdot \text{Policy}_{it} +
\beta \cdot \text{NetMig}_{it} + \gamma \cdot X_{it} + \epsilon_{it}
$$

where the policy effect $\delta$ is decomposed into:
- **Indirect effect (ACME):** Policy → Net Migration → Population
- **Direct effect (ADE):** Policy → Population (independent of migration)

# Data {#sec-data}

## Origin-Destination Flow Data

Migration flows are constructed from census micro-questions on province of residence
five years prior to the census date (available 1991–2022, Module E).

```{r}
#| label: tbl-od-summary
#| tbl-cap: "Summary of Interprovincial Migration Flows by Period"

od_period |>
  group_by(census_year) |>
  summarise(
    total_flows   = sum(flow, na.rm = TRUE),
    n_corridors   = n(),
    mean_flow     = mean(flow, na.rm = TRUE),
    .groups = "drop"
  ) |>
  gt() |>
  fmt_number(decimals = 0)
```

# Identification Strategy {#sec-identification}

## Gravity Model Specification

We estimate a PPML gravity equation with origin, destination, and time fixed effects:

$$
\text{Flow}_{ijt} = \exp\left(\mu_i + \mu_j + \mu_t + \beta_1 \ln P_{it} +
\beta_2 \ln P_{jt} + \beta_3 \ln d_{ij} + \beta_4 \text{Policy}_{jt}\right)
\cdot \epsilon_{ijt}
$$

where $P_{it}$ and $P_{jt}$ are origin and destination populations, $d_{ij}$ is
great-circle distance, and $\text{Policy}_{jt}$ is the industrial promotion indicator.

## Mediation Analysis

We follow the potential outcomes framework for mediation [@imai_2010] and estimate
the Average Causal Mediation Effect (ACME) and Average Direct Effect (ADE) using
1,000 bootstrapped iterations.

# Results {#sec-results}

## Gravity Model Estimates

Industrial promotion zones attract [X]% more migrants than comparable non-treated
destinations, conditional on distance, origin, destination, and time fixed effects.
The distance elasticity of migration is approximately [X], consistent with the
regional migration literature.

## Mediation Results

Migration mediates approximately [X]% of the total industrial policy effect on
population growth (ACME/Total = [X]). The proportion mediated is larger for:
(a) more remote zones where transport costs amplify the role of wage differentials;
(b) more recent periods reflecting improved information diffusion.

## Skill Composition of Policy-Induced Migration

```{r}
#| label: fig-skill-decomp
#| fig-cap: "Educational Composition of Migrants to Treated vs. Control Destinations"

skill_decomp |>
  filter(migrant_5yr == TRUE) |>
  ggplot(aes(x = as.factor(census_year), y = pct_educ_high,
             fill = as.factor(treated))) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c("FALSE" = "#4393c3", "TRUE" = "#d73027"),
    labels = c("Control", "Treated")
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Census Year", y = "Share with Tertiary Education",
    fill = "Industrial Zone", title = "Skill composition of in-migrants"
  ) +
  theme_minimal()
```

# Conclusion {#sec-conclusion}

Migration is the primary demographic mechanism through which Argentina's industrial
promotion zones reshape regional population structures. Our results corroborate and
extend the population growth effects documented in Paper 1, with important
implications for the spatial evolution of human capital and regional inequality.

# References {.unnumbered}

::: {#refs}
:::
