---
title: "Industrial Promotion Policies and Population Growth: Causal Evidence from Argentina"
subtitle: "A Staggered Difference-in-Differences Approach"
author:
  - name: "Author Name"
    affiliation: "Institution"
    email: "author@institution.edu"
    corresponding: true
date: last-modified
abstract: |
  This paper examines the causal effect of industrial promotion zones on subnational
  population growth in Argentina over the period 1970–2022. Exploiting the staggered
  adoption of industrial promotion legislation across Argentine departments, we apply
  the Callaway-Sant'Anna (2021) estimator to isolate the average treatment effect on
  the treated (ATT) while accounting for heterogeneous treatment timing. We complement
  this with generalized synthetic control methods (Xu 2017) and spatial econometric
  models to detect geographic spillovers. Our results suggest that industrial promotion
  policies generate [X] percentage-point increases in annual population growth rates
  within treated areas, with effects concentrated in [time period] and mediated
  primarily through [mechanism]. Spatial decomposition reveals significant positive
  spillovers to neighboring areas up to [X] km from zone boundaries.
keywords:
  - industrial policy
  - population growth
  - difference-in-differences
  - Argentina
  - spatial econometrics
  - place-based policy
format:
  pdf:
    documentclass: article
    classoption: [12pt, a4paper]
    number-sections: true
    colorlinks: true
    fig-cap-location: bottom
    tbl-cap-location: top
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{dcolumn}
        \usepackage{amsmath}
        \usepackage{float}
        \usepackage{setspace}
        \doublespacing
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
bibliography: references.bib
csl: apa.csl
execute:
  echo: false
  warning: false
  message: false
  cache: true
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(targets)
library(here)
library(gt)
library(knitr)
library(patchwork)

# Load pre-computed targets
did_effects       <- targets::tar_read(did_effects)
did_twfe          <- targets::tar_read(did_twfe)
spillover_decomp  <- targets::tar_read(spillover_decomposition)
spatial_morans    <- targets::tar_read(spatial_morans)
summary_tab       <- targets::tar_read(summary_table)
```

# Introduction {#sec-intro}

Place-based development policies occupy a contested position in the academic and policy
literature [@kline_2013; @neumark_2015]. On one hand, theory predicts that spatially
targeted industrial incentives can overcome coordination failures and catalyze agglomeration
economies in lagging regions [@krugman_1991; @moretti_2010]. On the other hand, critics
argue that such policies may merely relocate economic activity across space rather than
generate genuine growth, with uncertain and potentially negative consequences for
population distribution [@glaeser_ward_2009].

This paper contributes to resolving this debate by examining the demographic footprint of
industrial promotion zones in Argentina — a country that has maintained one of the
developing world's most extensive systems of geographically targeted industrial incentives
since the 1970s [@schvarzer_1987; @azpiazu_1988]. Unlike most existing evaluations,
which focus on firm-level outcomes or employment, we directly analyze **population dynamics**
as the ultimate measure of whether place-based policies succeed in reshaping the
geographic distribution of economic activity and human capital.

## Research Question

**Do industrial promotion zones alter subnational population growth trajectories in Argentina,
and if so, through what mechanisms?**

We decompose this question into three specific hypotheses:

1. **H1 (Growth effect):** Departments with active industrial promotion zones experience
   higher population growth rates than comparable non-treated departments.

2. **H2 (Persistence):** Treatment effects persist beyond the initial policy period,
   reflecting agglomeration spillovers.

3. **H3 (Spatial spillovers):** Population growth effects extend to neighboring departments,
   consistent with commuter and migration hinterland expansion.

## Contribution

This paper makes three contributions to the literature. **First**, we provide the first
causally credible estimates of the demographic consequences of Argentina's industrial
promotion system using modern staggered difference-in-differences methods
[@callaway_santanna_2021]. **Second**, we contribute to the broader place-based policy
evaluation literature by demonstrating how population dynamics can serve as a unified
outcome integrating multiple dimensions of regional development. **Third**, we generate
a methodological template for evaluating industrial policy in data-constrained
middle-income country contexts.

# Theoretical Framework {#sec-theory}

## Place-Based Industrial Policy and Population Dynamics

The theoretical link between industrial promotion and population growth operates through
two primary channels. The **direct employment channel** predicts that subsidized firm
formation increases local labor demand, attracting in-migration and retaining existing
residents [@moretti_2010; @suedekum_2006]. The **multiplier channel** predicts that
manufacturing employment generates downstream employment in non-tradeable services,
amplifying population effects beyond the initially targeted sector [@moretti_2010].

We situate this analysis within the regional growth and agglomeration literature
[@glaeser_1992; @combes_2008] and the demographic transition framework
[@bongaarts_bulatao_1999], which predicts that economic restructuring reshapes
not only population size but also age structure, household composition, and fertility.

## Industrial Policy in Argentina: Historical Context

Argentina's industrial promotion system emerged in the 1970s as an explicit tool for
reducing regional inequality and stimulating industrial development in lagging provinces.
The key legislative milestones include:

- **Ley 22.021 (1979):** Promoted industrial development in San Luis, La Rioja,
  Catamarca, and San Juan through tax exemptions and credit subsidies.
- **Ley 19.640 (1972):** Created a special economic regime for Tierra del Fuego and
  the South Atlantic islands.
- **Provincial promotion laws:** Various provinces enacted complementary legislation
  extending similar incentives to sub-provincial zones.

# Data {#sec-data}

## Census Data (Module A)

We use microdata from five Argentine population censuses: 1970, 1991, 2001, 2010, and
2022, harmonized to consistent geographic units (departments/partidos) using the
crosswalk described in the replication appendix. The harmonization procedure follows
@ipums_international and is detailed in [@sec-appendix-harmonization].

Key variables extracted include: population counts by age and sex, place of birth,
five-year residence (available from 1991), educational attainment, and labor force
participation.

## Industrial Promotion Zone Data (Module B)

Our primary treatment variable is derived from a georeferenced dataset of industrial
promotion zones, constructed from the legislative record. For each zone, we code:
the policy start year, end year (or right-censored if still active), geographic
boundary (polygon), incentive type (tax exemption, infrastructure, regulatory relief),
and primary targeted sector.

## Summary Statistics

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary Statistics by Treatment Status"

summary_tab |>
  gt() |>
  fmt_number(columns = where(is.numeric), decimals = 3) |>
  tab_spanner(label = "Control", columns = starts_with("control")) |>
  tab_spanner(label = "Treated", columns = starts_with("treated")) |>
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_row_groups()
  )
```

# Identification Strategy {#sec-identification}

## Research Design

Our primary identification strategy exploits the **staggered rollout** of industrial
promotion legislation: departments received treatment at different dates, allowing us
to use not-yet-treated and never-treated departments as counterfactuals for early-treated
departments.

> [!NOTE]
> **Identifying assumption (Parallel Trends):** In the absence of treatment, treated
> and control departments would have followed parallel population growth trajectories.
> We test this assumption via pre-treatment event-study estimates and the
> Rambachan-Roth (2023) sensitivity analysis.

We implement the @callaway_santanna_2021 estimator, which is robust to treatment
effect heterogeneity — a critical property given the diversity of industrial promotion
laws across provinces and time periods.

## Threats to Identification

**Selection into treatment:** Provinces that lobbied for industrial promotion may
differ systematically from non-treated provinces. We address this by: (a) controlling
for pre-treatment population levels and trends; (b) using doubly-robust estimation;
(c) presenting results separately for "never-treated" vs. "not-yet-treated" control
groups.

**Spatial spillovers:** If treatment effects spill over to neighboring departments,
our standard DiD estimates may be biased. We explicitly test for spillovers using
spatial econometric methods (@sec-spatial).

# Results {#sec-results}

## Main Effect on Population Growth

@fig-event-study presents the event-study estimates. We find that departments
receiving industrial promotion zones experience a statistically significant increase in
population growth rates beginning approximately [X] years after policy adoption. The
absence of significant pre-trends is consistent with the parallel trends assumption.

```{r}
#| label: fig-event-study
#| fig-cap: "Event Study: Effect of Industrial Policy on Log Population"
#| fig-height: 5

targets::tar_read(fig_event_study)
```

```{r}
#| label: tbl-did-main
#| tbl-cap: "Main DiD Estimates: Effect on Log Population"

tibble(
  Specification      = c("Callaway-Sant'Anna (Overall ATT)", 
                         "TWFE (benchmark)", 
                         "Callaway-Sant'Anna (Dynamic, post-treatment avg)"),
  Estimate           = c(did_effects$overall$overall.att,
                         coef(did_twfe)["treated"],
                         mean(did_effects$dynamic$att.egt[did_effects$dynamic$egt >= 0], na.rm = TRUE)),
  SE                 = c(did_effects$overall$overall.se, NA, NA),
  p_value            = c(did_effects$overall$overall.att / did_effects$overall$overall.se |> 
                           pnorm(lower.tail = FALSE) * 2, NA, NA)
) |>
  gt() |>
  fmt_number(decimals = 4)
```

## Spatial Spillovers {#sec-spatial}

We test for spatial autocorrelation in post-treatment population growth using Moran's I.
All post-1991 census rounds show significant positive spatial autocorrelation (Moran's I
ranging from [X] to [Y]), indicating that population growth is spatially clustered.

The Spatial Durbin Model decomposes the total treatment effect into direct and indirect
(spillover) components. Approximately [X]% of the total effect is attributable to
indirect effects on neighboring departments, consistent with commuter hinterland and
supply-chain linkage mechanisms.

## Robustness Checks

We conduct the following robustness checks (full results in the online appendix):

1. **Alternative control groups:** Using only "never-treated" departments as controls.
2. **Alternative outcome:** Annual growth rate (vs. log population level).
3. **Anticipation effects:** Allowing 1–2 census periods of anticipation.
4. **Geographic bandwidth:** Limiting to departments within 200 km of treated zones.
5. **Rambachan-Roth sensitivity:** Testing sensitivity of pre-trend test to violations
   of the no-anticipation assumption.

# Conclusion {#sec-conclusion}

This paper provides the first causally credible estimates of the demographic consequences
of Argentina's industrial promotion zone system. Our results demonstrate that:

1. Industrial promotion zones generate measurable, positive effects on local population
   growth with a lag of approximately [X] years.
2. Effects are geographically heterogeneous: [strong / weak / null] effects in
   [region/sector].
3. Significant positive spillovers extend [X] km beyond zone boundaries.
4. Effects appear [persistent / transitory], with [amplification / decay] observed
   [X] years post-adoption.

## Policy Implications

Our findings suggest that place-based industrial policy can effectively alter
population distribution at the subnational level. However, [caveats on efficiency,
equity, and long-run fiscal sustainability].

## Limitations and Future Research

This analysis is limited by: (a) the absence of continuous annual population
series between censuses; (b) limited data on within-zone firm dynamics and employment;
(c) the inability to separately identify migration and natural increase channels (see
Paper 2 of this series).

Future work should examine: heterogeneous treatment effects by sector and incentive
type; the mechanisms linking industrial employment to demographic outcomes; and the
comparative experience of Brazil and other Latin American economies.

# References {.unnumbered}

::: {#refs}
:::

# Appendix {#sec-appendix .unnumbered}

## A. Census Harmonization Procedure {#sec-appendix-harmonization}

Details of the crosswalk construction, geographic harmonization, and imputation
procedures are documented in `data/documentation/codebook.md` and the
`R/01_harmonization.R` script.

## B. Industrial Zone Dataset Construction

## C. Additional Robustness Results

## D. Spatial Diagnostics
