---
title: "Industrial Zones and Land-Use Transitions: Forest Footprints of Place-Based Policy in Argentina"
author:
  - name: "Author Name"
    affiliation: "Institution"
    corresponding: true
date: last-modified
abstract: |
  We examine whether industrial promotion zones in Argentina generate measurable
  land-use and forest cover transitions in their vicinities. Combining satellite-derived
  forest cover data (Hansen Global Forest Watch) with our georeferenced industrial zone
  dataset, we apply land-use change models, spatial regression, and a forest transition
  framework to evaluate the environmental footprint of place-based industrial policy.
  We find that zones associated with agro-industrial and resource extraction sectors
  generate significant forest cover loss within [X] km of zone boundaries, while
  manufacturing-focused zones show [null / mixed] forest effects. We adapt methods
  from Sharygin (2020) to estimate displacement of forest-dependent populations
  from areas adjacent to expanding industrial zones.
keywords:
  - land-use change
  - forest transition
  - industrial policy
  - Argentina
  - satellite data
  - Hansen GFW
  - environmental demography
format:
  pdf:
    documentclass: article
    classoption: [12pt, a4paper]
    number-sections: true
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{amsmath}
        \usepackage{float}
        \usepackage{setspace}
        \doublespacing
  html:
    toc: true
    number-sections: true
    theme: cosmo
bibliography: references.bib
csl: apa.csl
execute:
  echo: false
  warning: false
  message: false
  cache: true
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(targets)
library(here)
library(gt)
library(sf)
library(ggplot2)

spatial_panel  <- targets::tar_read(spatial_panel)
displacement   <- targets::tar_read(displacement_est)
```

# Introduction {#sec-intro}

The environmental consequences of industrial promotion policies represent a
largely overlooked dimension of place-based development evaluation. Industrial zones
can transform landscapes through multiple pathways: directly through land conversion
for infrastructure and production; indirectly through population-driven urban expansion;
and through induced agricultural intensification in supply-chain hinterlands.

This paper examines the **land-use and forest transition footprint** of Argentina's
industrial promotion system. We draw theoretically from the Forest Transition framework
[@mather_1992; @meyfroidt_lambin_2011], which predicts a non-linear relationship
between economic development and forest cover, and integrate Sleeter et al.'s (2017)
scenario-based land change methodology adapted to economic policy contexts.

A distinctive contribution is our adaptation of @sharygin_2020's displacement
estimation framework — originally developed for wildfire-affected populations — to
the context of forest-dependent communities adjacent to expanding industrial zones.

## Research Questions

1. Do industrial promotion zones trigger measurable land-use transitions in peri-zone areas?
2. What is the net forest cover change associated with industrial zone designation?
3. Are there measurable displacement effects on forest-adjacent populations?

# Theoretical Framework {#sec-theory}

## Forest Transition Theory

The Forest Transition (FT) hypothesis posits that as economies develop, forest cover
follows a U-shaped trajectory: initial deforestation gives way to reforestation as
agricultural productivity improvements and off-farm employment reduce pressure on
forest resources [@mather_1992]. Industrial promotion can accelerate or disrupt this
transition:

- **Acceleration pathway:** Manufacturing employment draws workers from forest-frontier
  agriculture, reducing deforestation pressure.
- **Disruption pathway:** Resource-extraction zones increase deforestation through
  direct land conversion and induced migration.

## Land-Use Change Modeling

We adapt the transition probability matrix approach of @sleeter_2017 to classify
land transitions across six categories:
```
(1) Forest → Urban       (deforestation + urbanization)
(2) Forest → Agriculture (agricultural expansion)
(3) Agriculture → Urban  (peri-urban expansion)
(4) Urban → Industrial   (zone-induced conversion)
(5) Stable Forest        (no transition)
(6) Other                (grassland, wetland, bare)
```

# Data {#sec-data}

## Forest Cover Data (Module C)

We use Hansen et al. (2013) Global Forest Change data at 30m resolution, available
2000–2022, to measure:
- Baseline tree cover (2000)
- Annual forest loss (2001–2022)
- Annual forest gain (2000–2022)

## Industrial Zone Data (Module B)

Described in Paper 1. For this analysis, we additionally code:
- Sector type (manufacturing, agro-industrial, resource extraction, mixed)
- Infrastructure footprint (road access, water, electricity)

# Identification Strategy {#sec-identification}

## Spatial Regression Discontinuity

Our primary identification strategy exploits the sharp geographic boundary of
industrial zone designations. Using a **Spatial Regression Discontinuity Design**
(Spatial RDD), we compare forest cover change rates just inside vs. just outside
zone boundaries, controlling for geographic characteristics and pre-designation trends.

## Difference-in-Differences (Land Cover)

We complement the spatial RDD with a department-level panel DiD, using departments
adjacent to — but not containing — industrial zones as a comparison group.

# Results {#sec-results}

## Forest Cover Transitions

```{r}
#| label: fig-forest-trend
#| fig-cap: "Forest Cover Change Near Industrial Zones (2000–2022)"

spatial_panel |>
  filter(!is.na(treecover_pct)) |>
  group_by(census_year, treated) |>
  summarise(mean_forest = mean(treecover_pct, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = census_year, y = mean_forest, color = as.factor(treated))) +
  geom_line(linewidth = 1) +
  geom_point() +
  scale_color_manual(
    values = c("FALSE" = "#4393c3", "TRUE" = "#d73027"),
    labels = c("Control Departments", "Industrial Zone Departments")
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(
    x = "Census Year", y = "Mean Tree Cover (%)",
    color = NULL, title = "Forest cover trends by treatment status"
  ) +
  theme_minimal()
```

## Displacement Estimation

Following @sharygin_2020, we estimate the population displaced from areas adjacent
to expanding industrial zones by comparing observed population with the synthetic
counterfactual from our generalized synthetic control model (Paper 1).

```{r}
#| label: tbl-displacement
#| tbl-cap: "Estimated Policy-Induced Displacement by Period"

displacement$total |>
  gt() |>
  fmt_number(columns = where(is.numeric), decimals = 0) |>
  cols_label(
    census_year = "Period",
    est_displaced_total = "Estimated Displaced (N)",
    n_depts_displaced   = "Departments with Deficit"
  )
```

# Conclusion {#sec-conclusion}

Industrial promotion zones in Argentina generate heterogeneous land-use and forest
transition effects that vary systematically by sector type: agro-industrial and
resource zones accelerate forest conversion, while manufacturing zones show smaller
and sometimes negative forest effects. Displacement of forest-adjacent populations
is a measurable consequence, concentrated in [regions/periods].

# References {.unnumbered}

::: {#refs}
:::
